Six steps to a better relationship with your future self.

### 2011-03-26 workflow.tex
Version given to Ashly. Perhaps the final version.

\usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\footnotesize}
\fvset{listparameters={\setlength{\topsep}{0pt}}}
\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
\SweaveOpts{keep.source=TRUE}  %Show comments
\fvset{fontsize=\footnotesize} %slightly smaller verbatim font

@@

<<initialize,echo=FALSE,results=hide>>=
  ##First, just setup the \R environment:
  options(SweaveHooks=list(
            fig=function(){
              par(mar=c(3.5, 3, 1.1, 0),
                  pty="s",
                  mgp=c(1.5,0.5,0),
                  oma=c(0,0,0,0))},#,
            echo=function(){options(continue=" ") ##Don't show "+" prompts,
                            options(prompt=" ")}
            ),
          digits=3,
          scipen=8,
          width=50,
          error=function(){options(prompt="> ",continue="+ ");NULL}
          )

@

@@

%%% This data frame will be used throughout the analyses.
%%% Putting it up top, so we can use it in the comments section.
<<read.data,echo=FALSE,results=hide>>=
##  This read.csv line is slightly more complex than needed because of
##  the particular encoding used for characters in the source file.
dem.nations.df<-read.csv(url("http://www.hks.harvard.edu/fs/pnorris/Data/Democracy%20CrossNational%20Data/Democracy%20Crossnational%20Data%20Spring%202009.csv",encoding="ISO_8859-1"),as.is=TRUE)
row.names(dem.nations.df)<-dem.nations.df$Natabrv ##make the row names more helpful
good.df<-na.omit(dem.nations.df[,c("Gini2004","protac2000","Nation","meanpr")])
good.df$gini04<-good.df$Gini2004/100
good.df$protac00<-good.df$protac2000
@

@@

\paragraph{Step 1} Know the provenance of results so that 
future  or current collaborators can quickly and easily reproduce
and thus change and improve upon the work.% --- to know what
% commands created which figures and tables.

@@

<<fig1code,echo=FALSE,fig=FALSE>>=
##  Make a scatterplot of Protest by Inequality
with(good.df,plot(gini04,protac00,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
                  cex=.8))

##  Label a few interesting points
with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
     text(gini04,protac00,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
@

\begin{figure}[H]
  \begin{center}
<<fig1plot,echo=FALSE,fig=TRUE,width=4,height=4>>=
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
<<fig1code>>
@
    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
    \label{fig:giniprot}
  \end{center}
\end{figure}

@@

<<maketable, echo=FALSE, results=hide>>=
  lm1<-lm(protac00~gini04+meanpr,data=good.df)  #  Run the regression

##  The next lines extract relevant and useful output from the
##  regression object, following Neal Beck's TPM article 2010.
themat<-cbind(summary(lm1)$coef[,c("Estimate","Std. Error")],confint(lm1))
n<-length(lm1$fitted.values)
r2<-signif(summary(lm1)$r.squared,2)
est.sigma<-signif(summary(lm1)$sigma,2)
colnames(themat)[3:4]<-c("~{}","~{}")
row.names(themat)<-c("Intercept","Income Inequality (lower=more equal)","Mean Political Rights (lower=more rights)")

library(xtable)
thextab<-xtable(themat,align=c("l",rep(".",4)),digits=1,
                caption="People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, adjusting for average political
  rights as measured by Freedom House 1980--2000. Data from \\citep{norris2009data}.",label="tab:protest")
@

<<printtable,echo=FALSE,results=tex>>=
  print(thextab,include.colnames=FALSE,
        hline.after=NULL,
        table.placement="H",
        add.to.row=list(
          pos=list(0,3),
          command=c(
            "\\toprule \n &
             \\multicolumn{1}{c}{Coef} &
             \\multicolumn{1}{c}{Std. Err.} &
             \\multicolumn{2}{c}{95\\% CI} \\\\ \\cmidrule(r){2-5} ",
            paste("& \\multicolumn{4}{c}{n: ",n,",",
                  "   resid.sd: ",signif(est.sigma),
                  ", R$^2$: ",r2,"} \\\\ \\bottomrule",sep="")
            )
          ))
@

@@

<<unit.test>>=
##  The test function:
test.times.2.fn <- function(){
  ##  This function tests times.2.fn
  if (times.2.fn(thenumber=4) ==  8 &
      times.2.fn(thenumber=-4) == -8) { 
    print("It works!")
  } else { print("It does not work!")
         }
}

##  The function:
times.2.fn <- function(thenumber){
  ##  This function multiplies a scalar number by 2
  ##  thenumber is a scalar number
  thenumber+2
}

##  Use the test function
test.times.2.fn()
@

@@

I have experimented with three systems so far: (1) for one paper we
simply included a Sweave document and data files into a compressed
archive \citep{bowers2005dataverse}; (2) for another more computing
intensive paper we assembled a set of files that enabled reproduction
of our results using the \texttt{make} system
\citep{bowers2008dataverse}; and (3) recently I have tried the
``compendium'' approach
\citep{gentleman2005reproducible,gentleman2007statistical} which
embeds an academic paper within the R package system
\citep{bowers2011dataverse}. The benefit of this last approach is that
one is not required to have access to a command line for \Verb+make+:
the compendium is downloadable from within R using
\Verb+install.packages()+ and is viewable using the \Verb+vignette()+
function in any operating system than runs R.\footnote{Notice that my
  reproduction archives and/or instructions for using them are hosted
  on the \href{http://thedata.org/book/learn-about-project}{Dataverse},
  which is another system to enhance academic collaboration across
  time and space.}  The idea that one ought to be able to install and
run and use an academic paper just as one installs and uses
statistical software packages is very attractive, and I anticipate that
it will become ever easier to turn papers into R packages as creative
and energetic folks turn their attention to the question of
reproducible research.

\paragraph{Step \thesection} We all learn by doing. When we share
reproduction materials we improve both cumulation of knowledge as
articulated in \citep{freese2007replication,king1995replication} and
our methods for doing social science.  As we copy and improve upon
each others modes of doing work we enhance our collective ability to
believe each other for future scholars to believe us, too.

@@

\end{document}
<<>>=
options(prompt="> ",continue="+ ")
@

>>>>

\usepackage[noae,nogin]{Sweave} %Declaring this explicitly so that I can modify the Sweave environment here rather in Sweave.sty which will be c
\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=.5em,fontsize=\footnotesize}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=.5em,fontsize=\footnotesize}
\DefineVerbatimEnvironment{Scode}{Verbatim}  {xleftmargin=.5em,fontsize=\footnotesize}
\fvset{listparameters={\setlength{\topsep}{0pt}}}
\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
  %Show comments
\fvset{fontsize=\footnotesize} %slightly smaller verbatim font

@@

@@

%%% This data frame will be used throughout the analyses.
%%% Putting it up top, so we can use it in the comments section.

@@

\paragraph{Step \thesection} If we know the provenance of results, 
future  or current collaborators can quickly and easily reproduce
and thus change and improve upon the work.

@@

\begin{figure}[H]
  \begin{center}
\includegraphics{workflow-fig1plot}
    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}
    \label{fig:giniprot}
  \end{center}
\end{figure}

@@

% latex table generated in R 2.12.2 by xtable 1.5-6 package
% Sat Mar 26 09:57:24 2011
\begin{table}[H]
\begin{center}
\begin{tabular}{l....}
   \toprule 
 &
             \multicolumn{1}{c}{Coef} &
             \multicolumn{1}{c}{Std. Err.} &
             \multicolumn{2}{c}{95\% CI} \\ \cmidrule(r){2-5} Intercept & 1.5 & 0.2 & 1.2 & 1.8 \\ 
  Income Inequality (lower=more equal) & -1.0 & 0.4 & -1.9 & -0.2 \\ 
  Mean Political Rights (lower=more rights) & -0.2 & 0.0 & -0.2 & -0.1 \\ 
   & \multicolumn{4}{c}{n: 53,   resid.sd: 0.28, R$^2$: 0.57} \\ \bottomrule\end{tabular}
\caption{People living in countries with unequal
  income distributions report less protest activity to World Values
  Survey interviewers than people living in countries with relatively
  more equal income distributions, adjusting for average political
  rights as measured by Freedom House 1980--2000. Data from \citep{norris2009data}.}
\label{tab:protest}
\end{center}
\end{table}

@@

\begin{Schunk}
\begin{Sinput}
 ##  The test function:
 test.times.2.fn <- function(){
   ##  This function tests times.2.fn
   if (times.2.fn(thenumber=4) ==  8 &
       times.2.fn(thenumber=-4) == -8) { 
     print("It works!")
   } else { print("It does not work!")
          }
 }
 ##  The function:
 times.2.fn <- function(thenumber){
   ##  This function multiplies a scalar number by 2
   ##  thenumber is a scalar number
   thenumber+2
 }
 ##  Use the test function
 test.times.2.fn()
\end{Sinput}
\begin{Soutput}
[1] "It does not work!"
\end{Soutput}
\end{Schunk}

@@

I have experimented with three systems so far: (1) for one paper we
simply included a Sweave document and data files into a compressed
archive \citep{bowers2005dataverse}; (2) for another more computing
intensive paper we assembled a set of files that enabled reproduction
of our results using the \texttt{make} system
\citep{bowers2008dataverse}; and (3) recently I have tried the
``compendium'' approach
\citep{gentleman2005reproducible,gentleman2007statistical} which
embeds an academic paper within the R package system
\citep{bowers2011dataverse}. The benefit of this last approach is that
one is not required to have access to a command line for \Verb+make+:
the compendium is downloadable from within R using
\Verb+install.packages()+ and is viewable using the \Verb+vignette()+
function in any operating system than runs R.\footnote{Notice that my
  reproduction archives and/or instructions for using them are hosted
  on the
  \href{http://thedata.org/book/learn-about-project}{Dataverse}, which
  is another system designed to enhance academic collaboration across
  time and space.}  The idea that one ought to be able to install and
run and use an academic paper just as one installs and uses
statistical software packages is very attractive, and I anticipate
that it will become ever easier to turn papers into R packages as
creative and energetic folks turn their attention to the question of
reproducible research.

\paragraph{Step \thesection} We all learn by doing. When we share
reproduction materials we improve both cumulation of knowledge and our
methods for doing social science
\citep{freese2007replication,king1995replication}.  As we copy and
improve upon each other's modes of doing work we enhance our collective
ability to believe each other and for future scholars to believe us, too.

@@

\end{document}
\begin{Schunk}
\begin{Sinput}
 options(prompt="> ",continue="+ ")
\end{Sinput}
\end{Schunk}
